Automated data-driven symbolic modeling of the lift forces acting on flapping wings

### 2021-02-07 paper/main.tex
Minor changes

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared". -->> Change to V^2
% everywhere (see Whitney Wood paper just below equation 2.18).
% TODO: If U-squared is "velocity squared" then  make sure that's consistent
% with the use of $v^2$ in the L_{trans} formula (equation 1) and elsewhere.
% TODO: Decide on better terms than "fitted" and "not fitted".
% TODO: Explain that the 5 EQ models were taken from the Pareto frontier of
% complexity vs. accuracy at the time of convergence.
% TODO: Explain that C_1,...,C_5 are not the same in each EQ. Maybe do C_i,_j
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(V^2\) Planar 1 & \(L_{planar},  F_{pendulum}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(V^2\) Planar 2 & \(L_{planar}, F_{pendulum}, F_{deviation}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(V^2\) Total & \(L_{trans}, F_{pendulum}, F_{deviation}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}, F_{pendulum}, F_{deviation}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans},
    AM_{trans-rot}, F_{pendulum}, F_{deviation}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}

@@

Inertial forces acting on the wings are modeled as a sum of two contributions.
The primary contribution results from the pendulum-like swinging motion that
occurs when the deflection angle ($\psi$) flips at stroke reversal.  The
secondary contribution results from small angle deviations from the horizontal
flapping plane ($\theta$). These inertial forces are given by:
\begin{equation}\label{sdbla}
F_{pendulum} = Y_{CM}M(\ddot{\psi}\sin\psi + \dot{\psi}^2\cos\psi)
\end{equation}
and
\begin{equation}\label{sdbla2}
F_{deviation} = -X_{CM}M(\ddot{\theta}\cos\theta - \dot{\theta}^2\sin\theta),
\end{equation}
\noindent where \(X_{CM}\) and \(Y_{CM}\) are the span-wise and chord-wise
locations of the center of wing area and \(M\) is the mass of the wing surface.
All of our analytical models for comparison include $F_{pendulum}$ and all
except $V^2$ Planar 1 include $F_{deviation}$.

% TODO: Remove all occurrences of the word "pareto" from the text, figures and filenames.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/mae_nofit}
\caption{Pareto plot of all data-driven (``EQ Models'') and analytical models
  using mean values for model coefficients. All data-driven models were smaller
  and more accurate than analytical models.\label{fig:pareto_mae}}
\end{figure}

>>>>

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared". -->> Change to V^2
% everywhere (see Whitney Wood paper just below equation 2.18).
% TODO: If U-squared is "velocity squared" then  make sure that's consistent
% with the use of $v^2$ in the L_{trans} formula (equation 1) and elsewhere.
% TODO: Decide on better terms than "fitted" and "not fitted".
% TODO: Explain that the 5 EQ models were taken from the Pareto frontier of
% complexity vs. accuracy at the time of convergence.
% TODO: Explain that C_1,...,C_5 are not the same in each EQ. Maybe do C_i,_j
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(V^2\) Planar 1 & \(L_{planar},  F_{pendulum}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(V^2\) Planar 2 & \(L_{planar}, F_{pendulum}, F_{deviation}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(V^2\) Total & \(L_{trans}, F_{pendulum}, F_{deviation}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}, F_{pendulum}, F_{deviation}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans},
    AM_{trans-rot}, F_{pendulum}, F_{deviation}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the 10 test
  experiments.}
\end{table}

@@

Inertial forces acting on the wings are modeled as a sum of two contributions.
The primary contribution results from the pendulum-like swinging motion that
occurs when the deflection angle ($\psi$) flips at stroke reversal.  The
secondary contribution results from small angle deviations ($\theta$) from the horizontal
flapping plane. These inertial forces are given by:
\begin{equation}\label{sdbla}
F_{pendulum} = Y_{CM}M(\ddot{\psi}\sin\psi + \dot{\psi}^2\cos\psi)
\end{equation}
and
\begin{equation}\label{sdbla2}
F_{deviation} = -X_{CM}M(\ddot{\theta}\cos\theta - \dot{\theta}^2\sin\theta),
\end{equation}
\noindent where \(X_{CM}\) and \(Y_{CM}\) are the span-wise and chord-wise
locations of the center of wing area and \(M\) is the mass of the wing surface.
All of our analytical models for comparison include $F_{pendulum}$, and all
except $V^2$ Planar 1 include $F_{deviation}$.

% TODO: Remove all occurrences of the word "pareto" from the text, figures and filenames.
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{figures/mae_nofit}
\caption{Pareto plot of all data-driven (``EQ Models'') and analytical models
  using mean values for model coefficients. All data-driven models were smaller
  and more accurate than analytical models.\label{fig:pareto_mae}}
\end{figure}
